Taking Dyads Seriously

### 2018-02-08 paper/4_app.tex
a few changes to the aps section

We choose five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}. These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade. It is worth noting, however, that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample. The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states (e.g., the USA).

These studies were also selected based on how recent they were and whether they had more then 100 citations. Each of these pieces was published in a prominent journal and is well-known in the literature. Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables. In addition, most of these studies adjust the standard errors in an attempt to account for the clustering of observations. We follow the strategy of the original study in our re-estimation of the original models.

@@

We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.} The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.

>>>>

We choose five recent studies from the broad field of international relations and international political economy that utilize relational data \citep{reiter:stam:2003, mcdonald:2004,  rose:2004, weeks:2012, gibler:2017}. These articles demonstrate the broad range of application for the AME framework, which include testing the effect of both dyadic and monadic variables on varied conflict behaviors and trade.  The broader goal in doing this, beyond introducing the use of the AME framework in an applied setting, is to examine the extent to which interdependencies within typical dyadic data alter previous conclusions about international relations from recent empirical studies using a framework that a priori assumes conditional independence.

These studies were also selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline. Of course, by the criteria of influence many other papers could have been considered for the replication. However, as decribed in the following sections, the AME model has specific data requirements.}  Each of these pieces was published in a prominent journal and is well-known in the literature. Each used the standard approach in political science, which is to employ some form of a general linearized regression that ignores dyadic interdependencies, except as they may reveal themselves in included variables. In addition, most of these studies adjust the standard errors in an attempt to account for the clustering of observations. We follow the strategy of the original study in our re-estimation of the original models.

@@

We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.}  It is worth noting that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample. The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.
